Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis

### 2024-01-24 outputs/paper/paper.qmd
Updating and commiting some spellcheck changes to paper.qmd

Although the Univerdity of Toronto emphasizes its ongoing commitment to student safety, the improvement of services, and raising awareness of assault and voilence, there is very little data and literature that exists on the matter. This paper aids to contribute to the investigation of violence in and around the Universities district.

@@

To examine the rate of assaults within and around the University of Toronto from 2014-2023, this paper has been organized into the following sections: Data, Results, Discussion, and Conclusion.  In the Data section, I discuss the nature of the spreadsheet obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data. The Results section highlights trends found during the analysis process, while the Discussion section further evaluates the trends and presents insight. Lastly, the Conclusion section summarizes the main findings from this paper.

@@

## Data
The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto]. The data set used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates]. Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].

###Source and Data Collection
This dataset includes the Crime Data by Neighbourhood. Counts are available for Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide and Shooting & Firearm Discharges. Data also includes the crime rate per 100,000 population calculated using the population estimates provided by Environics Analytics. Comparing to assault count, assault rate provides a fairer comparison of the crime over time by taking into account the change in population in the region. The data was last refreshed on Jan 11, 2024 and is refreshed annually. 
I conducted the first step of basic data cleaning to eliminate all counts and rates not applicable to this paper, so all I was left with was assault rates per neighbourhood for each year. I then refined the nieghbourhoods to include only those that encompass the University of Toronto and then its surrounding areas. The city of Toronto recognizes 158 unique neighbourhoods and of the 158, 13 have been included in the cleaned data.

@@

There are several limitations to this analysis. Firstly, this analysis relies solely on a single metric (assault rates), which provides a narrow view of campus safety. Further, while assault rates does quantifies the incidents occurring, it does not account for the severity or nature of the assaults (does not break the data down by type of assault or whether it was a chargable offense). The term 'assault' encompasses a broad range of actions. Without information on the type of assault, it is difficult to provide insight into policy change, social action, and campus safety.
# References

>>>>

Although the University of Toronto emphasizes its ongoing commitment to student safety, the improvement of services, and raising awareness of assault and violence, there is very little data and literature that exists on the matter. This paper aids to contribute to the investigation of assault violence in and around the University of Toronto.

@@

To examine the rate of assaults within and around the University of Toronto from 2014-2023, this paper has been organized into the following sections: Data, Results, Discussion, and Conclusion. In the Data section, I discuss the nature of the dataset obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data. The Results section highlights trends found during the analysis process, while the Discussion section further evaluates the trends and presents insight. Lastly, the Conclusion section summarizes the main findings from this paper.

@@

## Data
The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto]. The data set used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates]. Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `KableExtra` [@rKableExtra], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].

###Source and Data Collection
This dataset includes the Crime Data by Neighbourhood. Counts are available for Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide and Shooting & Firearm Discharges. Data also includes the crime rate per 100,000 population calculated using the population estimates provided by Environics Analytics. Compared to assault count, assault rate provides a fairer comparison of the crime over time by taking into account the change in population in the region. The data was last refreshed on Jan 11, 2024 and is refreshed annually.

I conducted the first step of basic data cleaning to eliminate all counts and rates not applicable to this paper, so all I was left with was assault rates per neighbourhood for each year. I then refined the nieghbourhoods to include only those that encompass the University of Toronto and then its surrounding areas. The city of Toronto recognizes 158 unique neighbourhoods and of the 158, 13 have been included in the cleaned data.

```{r}
THIS IS FOR MY PHOTO
```

@@

There are several limitations to this analysis. Firstly, this analysis relies solely on a single metric (assault rates), which provides a narrow view of campus safety. Further, while assault rates does quantifies the incidents occurring, it does not account for the severity or nature of the assaults (does not break the data down by type of assault or whether it was a chargable offense). The term 'assault' encompasses a broad range of actions. Without information on the type of assault, it is difficult to provide insight into policy change, social action, and campus safety.

# References
